---
title: "Number of Students per Code-Specific Presentation"
author: Chinomnso Ekwedike
output: html_notebook
---

This analysis computes the number of `id_pres` associated with each `pres_specific`.

Load the necessary libraries.

```{r library, echo=TRUE}
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggvis))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(treemapify))
```


Load the data set.
```{r data, echo=TRUE}
student_data = read_csv("data/Student_data.csv", show_col_types = FALSE)

```



Simple exploratory data analysis (EDA) 

```{r column_names, echo=TRUE}
colnames(student_data)
```


```{r dimension, echo=TRUE}
dim(student_data)
```


Compute the number of `id_pres` associated with each `pres_specific` 


Retrieve the necessary columns and check the new data frame.
```{r pres_columns, echo=TRUE}
# retrieve the columns
new_st_data = student_data %>% select(pres_specific, id_pres)

# check the data subset
n_row = nrow(new_st_data)
n_column = ncol(new_st_data)
missing = any(is.na(new_st_data))

# display the outputs on separate lines
cat(paste("Rows: ", n_row, "\nColumns: ", n_column, "\nMissing Rows: ", missing))
```


Count the number of students and convert the table to a data frame.
```{r count, echo=TRUE}
# convert table to data frame and rename columns
st_count = as.data.frame(table(new_st_data$pres_specific)) %>% rename(pres_specific = Var1, num_of_students = Freq)

# display the data frame
kable(st_count)
```


Compute and Include the percentage of students associated with each each code specific presentation.

```{r percentage, echo=TRUE}
#compute the percentage of each students within each code specific presentation
new_student_data = st_count %>% mutate(percentage = round(100 * num_of_students / n_row, 1))
kable(new_student_data)
#write_csv(new_student_data, "pres_specific_count.csv")
```


Graphical summaries of the analysis.

```{r number, echo=TRUE}
# plot pres_specific against number of students
new_student_data %>% ggvis(~pres_specific, ~num_of_students, fill = ~pres_specific) %>% layer_bars() %>% add_axis("x", title = "pres_specific", title_offset = 80, tick_padding = 3, properties = axis_props(labels = list(angle = -70, align = "right", fontSize = 10))) %>% add_axis("y", title = "num_of_students", title_offset = 50)
```



Display the percentage in a tree map. The tree map provides an alternative to pie chart given the number of rows (22) in the data frame.

```{r mosaic, fig.width=10, include=FALSE}
treemap::treemap(new_student_data, index = "pres_specific", vSize = "percentage", type = "index", fontsize.labels = 12, fontcolor.labels = "black", inflate.labels = F, title = "Percentage of Students Treemap", fontsize.title = 25, palette = "Reds")
```



```{r tree_mapping, echo=TRUE}
ggplot(new_student_data, aes(area = percentage, fill = pres_specific, label = paste(pres_specific, "\n", percentage, "%"))) + 
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 10) + 
  theme(legend.position = "none")
```











